---
title: "Data"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r set up, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom",plot.title = element_text(face = "bold")))
```

For this project, we mainly looked at the shooting incident data that we acquired from [NYC open data](https://opendata.cityofnewyork.us/data/). During the project, we also acquired data from other sources to help explain and analyze the findings we have from the shooting incident data.The details of the data and our cleaning methods will be elaborated below. 

## Data Sources

1. NYC Shooting Incident Data: 

[NYPD Shooting Incident Data (Historic)](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8)

This data set contains 23,575 shooting incidents that occurred during January 01, 2016 to December 31, 2020. The data set was originally provided by the Police Department (NYPD). For each incident, information on the incident date, time, location (specific to the longitude and latitude of the incident), the perpetrator and the victim's sex, race, and age group were provided. 


2. Unemployment Rate Data

[Unemployment Rate](https://fred.stlouisfed.org/series/NYUR)

We downloaded the monthly unemployment rate in New York City from January ,2010 to December, 2020 from FRED, [Federal Reserve Bank of St. Louis](https://fred.stlouisfed.org/series/NYUR). The original source of the data is [U.S. Bureau of Labor Statistics](https://www.bls.gov/sae/).


3. Population in each Police Precinct Data

We downloaded a cleaned data set on population in each police precinct data from this [website](https://johnkeefe.net/nyc-police-precinct-and-census-data). It's source is the [U.S Census Bureau](https://data.census.gov/cedsci/). 

4. NYC Population by age group

We downloaded a clean data set on population in each age group from this
[Projected Population 2010-2040 - Total By Age Groups](https://data.cityofnewyork.us/City-Government/Projected-Population-2010-2040-Total-By-Age-Groups/97pn-acdf)

## Data Cleaning and Tidying

1. Shooting Incident Data

The main data set we used is the Shooting Incident Data. After downloading our data, we cleaned the data by renamed a few variable with reasonable names, re-coded some variables to present clear information, and wrote out a csv file with the cleaned data as the resulting data set that we will be using in the project. 

```{r eval=FALSE}
data_hist = 
  read_csv("data/NYPD_Shooting_Incident_Data__Historic_.csv") %>%
  janitor::clean_names() %>%
  rename("date" = "occur_date",
         "time" = "occur_time") %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  separate(time, into = c("hour", "minute"), sep = ":") %>% 
  mutate(
          boro = str_to_lower(boro),
          location_desc = str_to_lower(location_desc),
          perp_race = str_to_lower(perp_race),
          vic_race = str_to_lower(vic_race),
          perp_age_group = str_replace(perp_age_group, "UNKNOWN","unknown"),
          perp_sex = recode(perp_sex, "M" = "male","F" = "female","U" = "unknown"), 
          vic_sex = recode(vic_sex, "M" = "male","F" = "female","U" = "unknown"),
          jurisdiction_code = case_when(
            jurisdiction_code == 0 ~ "patrol", 
            jurisdiction_code == 1 ~ "transit",
            jurisdiction_code == 2 ~ "housing")
          ) %>% 
  arrange(year,month,day,hour,minute)

write.csv(data_hist, ".\\data\\NYPD_Shooting_Incident_Data_Clean.csv", row.names=FALSE)
```

2. Unemployment Rate Data

We initially found data from the [New York State Department of Labor](https://dol.ny.gov/labor-statistics-new-york-city-region) and cleaned it for a resulting data set. However, this set only contains data from 2015 to 2020. 

```{r eval=FALSE}
library(readxl)

path = "data/unemployment/borough_labor_force.xlsx"

unemployment_df = bind_rows(path %>% 
  excel_sheets() %>% 
  set_names() %>% 
  map(read_excel, path = path)) %>% 
  janitor::clean_names() %>% 
  drop_na() %>%
  subset(month != "Avg") %>% 
  rename("boro" = "area","unemployment_rate" ="unemployment_percent") %>% 
  mutate( boro = case_when(
    str_starts(boro,"B") ~ "bronx",
    str_starts(boro,"K") ~ "brooklyn",
    str_starts(boro,"Q") ~ "queens",
    str_starts(boro,"R") ~ "staten island",
    str_starts(boro,"N") ~ "manhattan"),
    month = str_to_lower(month),
    unemployment_rate = round(unemployment_rate/100,3)
    )

write.csv(unemployment_df, "data/unemployment/unemployment_cleaned.csv")
```

Therefore, we went back and looked for another source. Eventually, we used data from the[Federal Reserve Bank of St. Louis](https://fred.stlouisfed.org/series/NYUR), which was a pretty cleaned data set to start with. We also wrote out a resulting csv file for our use in the project as well. 

```{r eval=FALSE}
unemployed_complete= read_csv("data/unemployment/NYUR.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "date"), sep = "-") %>% 
  rename("unemployment_rate" = "nyur") %>% 
  mutate(unemployment_rate = round(unemployment_rate/100,3)) %>% 
  select(-date)

write.csv(unemployed_complete, "data/unemployment/unemployed_complete.csv")
```

3. Population in each Police Precinct 

Since the data set is already cleaned when we found it, we didn't have to do much tidying. However, the columns were cryptic. Therefore, we followed the code listed in the a Census Bureau PDF starting on this [page](https://www.documentcloud.org/documents/87708-pl94-171-1.html#document/p64), to re-code the column that we need to `total_pop`, which gives us the total population for each police precinct. 

```{r eval=FALSE}
percinct_pop = read.csv("data/nyc_2010pop_2020precincts.csv") %>%
  select(precinct_2020, P0010001) %>% 
  rename("total_pop" = "P0010001")
```

4. Population by age group

The data set [Projected Population 2010-2040 - Total By Age Groups](https://data.cityofnewyork.us/City-Government/Projected-Population-2010-2040-Total-By-Age-Groups/97pn-acdf) is already cleaned , but the age group interval are smaller than we used in analysis. Therefore, we add the population group from this data set together to fit the larger age group interval in our analysis.

```{r eval=FALSE}
age_census = read.csv("data/population_nyc.csv")%>%
  janitor::clean_names()%>%
  filter(borough == "NYC Total")%>%
  select(borough, age, x2010)

age_clean_census = data.frame("age_group"= c("0-18","19-24","25-44","45-64","65+"), "population" = c(age_census$x2010[1]+age_census$x2010[2], age_census$x2010[3],age_census$x2010[4]+age_census$x2010[5]+age_census$x2010[6]+age_census$x2010[7],age_census$x2010[8]+age_census$x2010[9]+age_census$x2010[10]+age_census$x2010[11],age_census$x2010[12]+age_census$x2010[13]+age_census$x2010[14]+age_census$x2010[15]+age_census$x2010[16]))%>%
  mutate(population = as.integer(population))


```

## Data description

**Shooting Incident Data**

The cleaned resulting data set contains 23,585 incidents and 22 variables. Some of the variables that we are interested in are:

  * `year` Year of the incident
  * `month` Month of the incident 
  * `day` Day of the incident
  * `hour` Time(hour) of the incident
  * `boro` The borough where the incident happens
  * `statistical_murder_flag` the outcome of the incident. TRUE means that the victim dies from the shooting, and FALSE means the opposite
  * `perp_age_group` Perpetrator's age group
  * `perp_sex` Perpetrator's sex
  * `perp_race`Perpetrator's race 
  * `latitude` Latitude of the incident
  * `longitude` Longitude of the incident. 
  
**Other Data**

We used the `unemployment rate` , `population by age group` and `police precint` data to further explore and analyze the findings that we obtained from our main `shooting incident` data. They are used in both the descriptive and statistical analysis in our project. `unemployment rate` has the monthly unemployment rate of NYC from 2010 to 2020.`police precinct` has the total population of each police precinct in NYC. 
  
