---
title: "Statistical analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(message = FALSE,warning = FALSE
)
```

# Goal
After visualization of the shooting incidences data in section descriptive analysis, we will go to statistical analysis to answer several questions that most of the nyc residents might care about.

In this section we will analysis the shooting case by the factor of the shooting, and carry out statistical analysis and data visualization about the shooting in New York City. We will asking some questions, and answer them using facts and plots.

#  Kruskal-Wallis analysis
## Which boro is most dangerous?

Due to recent raise in shooting events around the US, people become more conscious about their safety. In particular, students of Columbia university want to know whether the community around the university is comparably safe or danger in New York. Here we use bar plot to show the total shooting events from 2006 to 2020 in all 5 boros of New York: "bronx", "manhattan", "queens", "brooklyn" and"staten island".

```{r}
library(tidyverse)
library(dplyr)
library(forcats)
library(readr)
library(ggplot2)
library(plotly)
library(lubridate)

toDate <- function(year, month, day) {
    ISOdate(year, month, day)
}

data_hist = 
  read_csv("data/NYPD_Shooting_Incident_Data_Clean.csv") %>%
  mutate(
    statistical_murder_flag = as.character(statistical_murder_flag),
    statistical_murder_flag = recode(statistical_murder_flag, "FALSE" = "survived","TRUE" = "murdered")
  )
library(dplyr)
data = 
  data_hist %>%
  dplyr::select(month, day, year, hour, boro,precinct,
         location_desc, perp_age_group,
         perp_sex, perp_race) %>% 
  mutate(date = toDate(year, month, day)) %>%
  mutate(weekday = wday(date, label = TRUE, abbr = TRUE)) %>%
  mutate(COVID = date >= toDate(2020, 4, 1))
  
#unique(data_hist$boro)

```

```{r}
p1 = 
  data %>%
  count(boro) %>%
  mutate(
    proportion = round(n / sum(n), 2) 
  )  %>%
  top_n(n = 10, wt = n) %>%
  mutate(
    District = fct_reorder(as.factor(boro), n, .desc = TRUE)
  ) %>%
  ggplot(aes(x = District, y = n, fill = District)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none") +
  labs(
    
    x = "",
    y = ""
  )

ll1 = data %>%
  count(boro) %>%
  mutate(
    proportion = round(n / sum(n), 2) 
  )  %>%
  top_n(n = 10, wt = n) %>%
  mutate(
    District = fct_reorder(as.factor(boro), n, .desc = TRUE)
  )

ll1$event_by_thousand_pop = ll1$n / c(1472654,
        2736074, 1694251, 2405464, 495747) * 1000

p2 = ll1 %>% 
  top_n(n = 10, wt = event_by_thousand_pop) %>%
  ggplot(aes(x = District, y = event_by_thousand_pop, fill = District)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none") +
  labs(
    x = "",
    y = ""
  )


subplot(p1, p2, nrows = 1) %>% 
  plotly::layout(title = 'Bar plot of shooting by districts <br><sup>left: shooting , right: shooting / population (thousands) </sup>',
                 autosize = T)
```

In the bar plot above we realize that there are most shooting events in brooklyn, and second is bronx, where manhattan is only around forth of the shooting events in brooklyn. It seems ok. From the shooting events per thousand population bar chart, the highest is bronx and second is brooklyn, manhattan is significantly lower than those 2, so compared to those 2 boros, the manhattan district is relatively safer, that's a good news for all students in our university.

Now we wanna use Kruskal-Wallis test to statistically test whether some districts are more dangerous than others. We will use the precincts shooting per thousand people as the sample, and apply Kruskal-Wallis test on that sample. The reason why we don's use ANOVA but use the non parametric analysis is because the data is not normally distributed.

```{r}

precinct_pop = read.csv("data/nyc_2010pop_2020precincts.csv") %>%
  rename("total_pop" = "P0010001") %>% 
  dplyr::select(c(precinct_2020,total_pop))

data1 = data %>%
  count(boro, precinct) %>%
  left_join(precinct_pop, 
            by = c("precinct" = "precinct_2020"))

data1$shoot_per_thousand = data1$n / data1$total_pop * 1000

library("ggpubr")
p1 = ggline(data1, x = "boro", y = "shoot_per_thousand", 
       add = c("mean_se", "jitter"), 
       ylab = "shoot per thousand", xlab = "boros") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none")
ggplotly(p1, tooltip = c("x", "y")) 
```

Then we perform the kruskal.test.

```{r}
kruskal.test(shoot_per_thousand ~ boro, data = data1)
```

As the p-value is less than the significance level 0.05, we can conclude that there are significant differences in shooting rate between the boros. From the output of the Kruskal-Wallis test, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different. So we use the pairwise wilcox test to calculate pairwise comparisons between boros with corrections for multiple testing.

```{r}
pairwise.wilcox.test(data1$shoot_per_thousand, data1$boro,
                 p.adjust.method = "BH")
```

Under the 0.05 significance level, we know that there is a significant difference between shooting rate of manhattan and bronx, queens and bronx. And from the bar chart we know that difference means manhattan and queens are significantly safer than the bronx boro.

# Visualization of Shooting time and places
## Where do the shooting most happens?
Where do the shooting most happens? (The last bar is not applicable) From the plot we know that first 3 places are apartment and houses, and of the public area, the first 3 are grocery stor, bar / night club and comercial buildings.

```{r}
p1 = 
  data %>%
  count(location_desc) %>%
  mutate(
    proportion = round(n / sum(n), 2) 
  )  %>%
  top_n(n = 10, wt = n) %>%
  mutate(
    location = fct_reorder(as.factor(location_desc), n, .desc = TRUE)
  ) %>%
  ggplot(aes(x = location, y = proportion)) +
  geom_bar(stat = "identity")+
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), legend.position = "none") +
  labs(
    title = "Top Ten Places to occur shooting",
    x = "",
    y = ""
  )

ggplotly(p1, tooltip = c("x", "y")) 
```

## When do most of the shooting happens?

```{r}
heat_maps = function(dataset, title) {
  pre = 
  dataset %>%
  filter(COVID == FALSE) %>%
  plot_ly(
    x = ~hour, y = ~weekday, z = ~shooting, type = "heatmap", colors = "BuPu"
  ) %>%
  colorbar(title = "Pre-COVID", x = 1, y = 1) 

covid = 
  dataset %>%
  filter(COVID == TRUE) %>%
  plot_ly(
    x = ~hour, y = ~weekday, z = ~shooting, type = "heatmap", colors = "YlGn"
  ) %>%
  colorbar(title = "COVID", x = 1, y = 0.5) 

subplot(pre, covid, nrows = 2, margin = .05) %>%
  layout(title = title)
}

d2 = 
  data %>%
  group_by(COVID, hour, weekday) %>%
  summarise(
    shooting = n()
  )  

heat_maps(dataset = d2, title = "Heat Maps of Shootings")
```

From both Pre-COVID and COVID time, the shooting are most centered at night, especially 23 PM - 4 AM. For the Pre-COVID time, most of the shootings are in the midnight of weekends, and for COVID, it seems to be more dispersed in nights.

## Which age group of people is the most dangerous?

From the the visualization in the section of descriptive, we see that age group 18-24 and age group 25-44 have the highest shooting incidences. However, more statistical analysis are required to indicate if the age group 18-24 and group 25-44 are more dangerous or those two groups have larger proportion in the population.

```{r include=FALSE}
age_census = read.csv("data/population_nyc.csv")%>%
  janitor::clean_names()%>%
  filter(borough == "NYC Total")%>%
  select(borough, age, x2010)
 
```



```{r include=FALSE}
age_clean_census = data.frame("age_group"= c("0-18","19-24","25-44","45-64","65+"), "population" = c(age_census$x2010[1]+age_census$x2010[2], age_census$x2010[3],age_census$x2010[4]+age_census$x2010[5]+age_census$x2010[6]+age_census$x2010[7],age_census$x2010[8]+age_census$x2010[9]+age_census$x2010[10]+age_census$x2010[11],age_census$x2010[12]+age_census$x2010[13]+age_census$x2010[14]+age_census$x2010[15]+age_census$x2010[16]))%>%
  mutate(population = as.integer(population))


```



```{r }
vic_age_data = 
  data_hist %>% 
  filter(year >= 2006,
         vic_age_group != "NA",
         vic_age_group != "UNKNOWN"
         ) %>% 
  group_by(vic_age_group) %>%
  summarise(incident_case = n()) %>%
  mutate(
    prop = incident_case/age_clean_census$population * 100000 /15 ) %>%
  
  ggplot(aes(vic_age_group, y = prop))  + geom_col() +
  labs(x = "Victim age group",
       y = "Incident rate  per 100,000 people",
       title = "Incident rate per 100,000 people vs Victim age group"
       )

vic_age_data
```

The age group 18-24 have the highest incidents rate which 92.7 shooting incidents per 100,000 persons a year from 2006 to 2020 in New york city. The age 65+ have the lowest incidents rate which 1.02 shooting incidents per 100,000 persons a year from 2006 to 2020 in New york city. The age group least than 18 , 25-44, 45-64 have incidents rate of  15.9, 26.7, 5.10 shooting incidents per 100,000 persons a year from 2006 to 2020 in New York city. This statistical analysis indicates that age 18-24 is the most dangerous group in nyc. This results might be affected by behavior factor, such as, young people are more likely to do more out-door activity than children and older people, or people are also more aggressive in their 20s.



