---
title: "Shiny Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
```

Column {.sidebar}
-----------------------------------------------------------------------

```{r}

```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

```

### Chart C

```{r}

```
This is very similar to the usual flexdashboard template, with some small changes. First, we’ve added runtime: shiny to the YAML header. Second, we are using slightly different layout because we have a sidebar column which will hold our input widgets.

I’ll add the following libraries to the template, but we won’t need much else.

library(tidyverse)
library(viridis)
## Loading required package: viridisLite
library(p8105.datasets)

library(plotly)
## 
## Attaching package: 'plotly'
## The following object is masked from 'package:ggplot2':
## 
##     last_plot
## The following object is masked from 'package:stats':
## 
##     filter
## The following object is masked from 'package:graphics':
## 
##     layout